---
title: "Lab 03"
author: "farha hussaini"
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(sf)
``` 

# Data

```{r load-data, message = FALSE}
fl_votes <- st_read("data/fl_votes.shp", quiet = TRUE)
fl_votes %>%
  slice(1:6)
```


# Exercise 1

```{r add-winner}
fl_votes <- fl_votes %>%
mutate(winner20 = if_else(rep20 > dem20, "Republican", "Democratic"))
```


# Exercise 2

```{r fl-plot-1}
fl_votes %>% ggplot() +
geom_sf(data = fl_votes, aes(fill = winner20)) +
scale_fill_manual(values = c("#0015BC" , "#DE0100")) +
ggtitle("Florida's 2020 U.S. Presidential Election Results by County") +
theme_void()


```

# Exercise #3

```{r fl-props}
fl_votes <- fl_votes %>%
mutate(total16 = dem16 + rep16,
total20 = dem20 + rep20,
prop_rep16 = rep16 / total16,
prop_rep20 = rep20 / total20)
```

# Exercise 4

```{r fl-plot-2}

ggplot(fl_votes) +
geom_sf(aes(fill=prop_rep20)) +
scale_fill_gradient(low="#0015BC", high="#DE0100") +
  labs(title="Election 2020 Results", fill="Republican share of votes") +
theme_bw()


```

# Exercise 5

```{r fl-change}
fl_votes <- fl_votes %>%
mutate(diff_rep = prop_rep20 - prop_rep16)
```

# Exercise 6

```{r fl-plot-3}
ggplot(fl_votes) +
geom_sf(aes(fill=diff_rep)) +
scale_fill_gradient2(low="#0015BC", midpoint=0,
high="#DE0100") +
labs(title="Election 2020 Results", fill="change in
Republican share") +
theme_bw()
```

# Exercise 7


First graph tells us that while there are some blue counties in Florida, the state overall is dominated by red
counties, indicating a Republican stronghold.
The second visualization shows us that there are some counties in Florida where the Republican share of the
vote is lower, means that these counties may be more politically diverse or leaning towards the Democratic
party.
The third graph shows that there were some counties where the Republican share of the vote decreased from
2016 to 2020, indicating that these counties may be shifting away from the Republican party.
these visualizations have limitations. They do not take into account demographic factors, that may influence
voting patterns. Additionally, they only provide a snapshot of a specific point in time. Also, it is only
show the Republican and Democratic parties, and do not account for third-party candidates or independent
voters.

